---
title: "Strawberry EDA Report"
author: "Wuge Li"
output: pdf_document
date: "2023-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```

Introduction:

The data set I used for this report is strawberry data. 


The column names normally represent what this data will show you. So, first, 
Let's take a look at the column names of strawberry raw data

```{r}
strawberry <- read.csv("strawberry.csv", header = T)
colnames(strawberry)
```

From the column name of the data we know, the information about Program
(census or survey), year, period, state(origin of the strawberries) and other 
information(whether or not use chemical? Which chemical used when the strawberry
grow up)

After watching the data of strawberry. I am curious about what influence the quality
of strawberry? Year? Origin? Chemicals? Also, the relationship between the origin 
of strawberry produce and the market the strawberry sold. And whether the strawberry
fresh of not.





Data cleaning and tidying:

Look at origin of the strawberry
```{r}
state_all <- strawberry |> 
  group_by(State) |> 
  count()
state_all

state_max <- state_all$State[which(state_all$n ==  max(state_all$n))]
state_max
```
```{r}
ggplot(data = state_all, mapping = aes(reorder(State, desc(n)), n)) +
  geom_bar(stat = 'identity')+
  labs(title = "Count of strawberries in each states",
       x = "States",
       y = "Count") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
The graph above shows that the count of strawberries in each states by decreasing
order


From the chart and graph above, we know that California contributes the most strawberry


We know that California contributes the most strawberry, then we concentrate on
strawberry which origin is California 
```{r}
calif_census <- strawberry |> filter((State=="CALIFORNIA") & (Program=="CENSUS"))
calif_survey <- strawberry |> filter((State=="CALIFORNIA") & (Program=="SURVEY"))
nrow(calif_census)
nrow(calif_survey)
calif_cen_per <- nrow(calif_census)/(nrow(calif_survey)+nrow(calif_census))
calif_svy_per <- nrow(calif_survey)/(nrow(calif_survey)+nrow(calif_census))
calif_cen_per
calif_svy_per
```
```{r}
slices <- c(calif_cen_per, calif_svy_per)
labels <- c("CENSES", "SURVEY")
calif_percentages <- round(c(calif_cen_per * 100, calif_svy_per * 100), digits = 2)
calif_labels_with_percentages <- paste(labels, "\n", calif_percentages, "%")
pie(slices, labels = calif_labels_with_percentages, main = "CENSUS vs SURVEY(California)")
```

From above, we know that there are 30 programs of 1886 are CENSUS, and 1856 of
1886 are SURVEY. CENSUS Program takes 1.59% of whole Programs and SURVEY Program
takes 98.41% of whole Programs

Then I am curious about the number of CENSUS Programs and SURVEY Programs.
```{r}
strwb_census <- strawberry |> filter(Program == "CENSUS")
strwb_survey <- strawberry |> filter(Program == "SURVEY")

nrow(strwb_census)
nrow(strwb_survey)
strwb_cen_per <- nrow(strwb_census)/(nrow(strwb_survey)+nrow(strwb_census))
strwb_svy_per <- nrow(strwb_survey)/(nrow(strwb_survey)+nrow(strwb_census))
strwb_cen_per
strwb_svy_per
```
```{r}
slices <- c(strwb_cen_per, strwb_svy_per)
labels <- c("CENSES", "SURVEY")
strwb_percentages <- round(c(strwb_cen_per * 100, strwb_svy_per * 100), digits = 2)
strwb_labels_with_percentages <- paste(labels, "\n", strwb_percentages, "%")
pie(slices, labels = strwb_labels_with_percentages, main = "CENSUS vs SURVEY(Strawberry)")
```

However, when we looked at the programs of all strawberries, we found that CENSUS
Program takes 20.03% and SURVEY Programs take 79.97% of the whole programs.


Conclusion:

Based on the information and tidy data above, we can know that California and Florida
contributes much more strawberries than other states. SURVEY Programs contributes nearly
80% for the whole strawberries and contributes nearly 98% for California strawberries.

